Modelling long-term effects of management strategies in mixed fisheries

### 2019-04-11 doc/MSE_MAR_18.tex
version ready to be checked by Adriaan

Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged. However, individual quota cobined with the stochastic fisheries catches resulted in lower than intended exploitation rates. The TACs were set so that the total TACs corresponded to the harvest rate that resulted in maximum sustainable yields. The TACs were divided in individual quotas for each fishing vessel. The lower than intended harvest rates were caused by the fact that despite their planning, fishers often exhausted one of the quotas before the other. If discarding is not allowed, then individual fishers must then stop fishing. At the population levels of the two species this resulted in overal cacthes that were lower than the TACs. Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere \cite{Kempf2016, Rindorf2017, Hilborn2015}, but is generall attributed to differences in life-histories among fished species. In this study, the discrepancy between the intended annual quotas at MSY and realized catches is caused only by schochasticity in the system. This stochasticy in catch rates is observed in most fisheries  \cite{Sampson1988, VanOostenbrugge2004, Smith1980, Bernasconi2015}. It is caused by the complex migration biology and the effects of environmental factors on fish behaviour. A well functioning quota market would allow vessels to swap excess quota for constraining and thus bring realized catches closer to quotas.

The model aimed to evaluate the biological and economic effects of adopting a major policy change such as the CFP \cite{CFP2013}. Starting from an overfished mixed fishery under open access conditions but with a landing obligation, the response of individual effort allocation and the interannual population dynamics are explicitly modeled, while adopting MSY objectives. The analysis extended earlier approaches that used short-term state-dependent decision-making models (e.g. \cite{Alzorriz2018, Batsleer2016, Poos2010}) by including interannual population dynamics of fish stocks, and a management procedure that mimics the decisions of a management system. Such an MSE framework was thought of as a minimum realistic model \cite{Kell2007, Punt1995}. To keep the outcomes of the model simple, and to show the implications of the fishing effort allocation, the two fish populations modelled here are exactly the same apart from their spatial distribution difference (Fig. \ref{f:distribution}). As a result, if the landing obligation changes selectivity patterns in the short-term as predicted in \cite{Alzorriz2018, Batsleer2016}, the present MSE modeling approach allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling how changes in exploitation rates impact fish populations and yields. ERROR IN DECISION MAKING, NO DISCARING ALLOWED BY FISHERS, AND PERFECT COMPLIANCE TO THIS

The forecasted effects of fisheries management in mixed fisheries result from spatial heterogeneity in resources and fishers adaptive behaviour. This emphasizes that human behaviour and decision-making are intrinsic parts of conservation and natural resource management and the need for incorporating detailed fishing dynamics in management strategy evaluation. Historically, the focus of management strategy evaluation has been primarily on population dynamics in single and multispecies fisheries \cite{Butterworth2007, Dichmont2006b, Punt1999, Punt2009, Smith1999, Punt2016}. Explicitly incorporating direct effects of human behaviour of those who harvest stocks is equally important \cite{Fulton2011,Milner-Gulland2012}, but relatively few studies have attempted this \cite{Andersen2010, Ono2018}. Our MSE that links a dynamic state variable model for fisher behavior to a multi-species age structured model provides an alternative to those approaches.

When the studied mixed fishery was unmanaged, with unconstrained quota for both species (pre-manage period) and no discarding allowance, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$). Overfishing occurs in both stocks when fishers exceed $H_{max}$, whereas rebuilding requires reducing exploitation below $H_{max}$ (Fig. \ref{f:yields}). Therefore, during the pre-manage period, a high exploitation rate caused a decline in total biomass and fishers mainly harvest the individuals that would normally be added to the population, with younger fish being targeted for. Our results suggested that given the spatial and price species distribution parameters (Fig. \ref{f:distribution} and \ref{f:prices}), fishing in both the northern and southern fishing grounds were equally profitable; thus, the distribution of fishing effort was equally distributed (Fig. \ref{f:effort}).

Moving from an unmanaged period to a situation where one or both species were quota regulated by implementing MSY objectives while maintaining a LO management regime (i.e. main CFP objectives: exploiting at MSY and no discarding allowance) led to a very restrictive fishing activity (Fig. \ref{f:effort}). Fishing effort was constrained by one or two quota stocks, as predicted in general terms by \cite{Ulrich2017}, resulting in drastic reductions of effort over the short-term and mid-term. In addition to reconciling fishery and conservation objectives (such as the CFP), setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig. \ref{f:effort}). However, as \cite{Prellezo2016a} concluded, since effort was drastically reduced to achieve the objectives of the CFP; the increase in net revenue gains were only experienced by some vessels. Consequently, a significant part of the fleet had to cease its activity (Fig. \ref{f:effort}).

Nevertheless, the framework used made several assumptions: that there was perfect knowledge of the stochastic nature of catch rates, full compliance of fishers with the discarding policy, lack of transferability of individual quotas, and lack of interference competition among fishers \cite{Alzorriz2018}. IMPORTANT< UTILITY FUNCTION IS ANNUAL NET REV, NO SOCIAL DYNAMICS. Regarding fish populations, also important assumptions were established: negligible natural mortality, constant recruitment, constant growth, constant migration, and constant size-dependent prices. Individual choices on fishing behaviour were assumed to be made once a month. However, despite these assumptions, the minimum realistic model developed in this study allowed us to simulate the dynamics of a mixed fishery and examine the effects of single species quota management.

In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of any discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annaul quota swaps. Allowing the latter could potentially remedy the underutilisation of quota in the two quota situation, given that this reuslts from the stochsticity in catches, which are uncorrelated.

 We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY. However, applying any kind of exemption or flexibility beyond the MSY would be negative \cite{Prellezo2016}. Consequently, a critical feature of management bodies is to ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation. Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries \cite{Alzorriz2018}. The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance. With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Batsleer2013}. In this MSE framework, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations. However, if detection rates are low and the short-term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance. IF YOU WANT TO STUDY THIS, THEN DETAILED PARAMETRIZATION IS NEEDED THAT INCLUDES CASE DEPENDENT LH CHARATRISTICS AND MIGRATIONS
 %Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation. The results highlighted the importance of accounting for technical interactions and their temporal dynamics in both quota allocation and fleet dynamics to design a realistic multispecies fishery management strategy analysis

Finally, a dynamic state variable model was used as a tool to create a quick quota allocation and fleet dynamics model. Many detailed fleet dynamics models have been developed over the past decades \cite{Andersen2010, Ono2018, Simons2015a} and they could have been used to include spatial factors as another dimension in the MSE. However, following \cite{Dowling2011}, incorporating uncertainty in errors in human dynamic decision-making help us to understant and predict changes in human behaviour in response of adopting a major policy change such as the CFP \cite{CFP2013}. The effect of introducing a realistic multispecies fishery within this MSE framework still remains the next step to be taken. Modelling and predicting realistic human behavior with its associated uncertainty in an ever-changing socioecological system will remain a major challenge to address in the future \cite{Ono2018}.

@@

%Although this may seem implausible too, a drastic reduction of effort would increase profits if MSY and LO were contemplated. However, no flexibilities or year transfer swaps were consider. Be carefull if applying flexibilities, because as Prellezo said, any increased in mortality beyond the MSY would be negative...
%prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative. Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future. This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.

%The largest yield (or catch) that can be taken from a species' stock over an indefinite period. Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely. Under the assumption of logistic growth, resource limitation does not constrain individuals’ reproductive rates when populations are small, but because there are few individuals, the overall yield is small. At intermediate population densities, also represented by half the carrying capacity, individuals are able to breed to their maximum rate. At this point, called the maximum sustainable yield, there is a surplus of individuals that can be harvested because growth of the population is at its maximum point due to the large number of reproducing individuals. Above this point, density dependent factors increasingly limit breeding until the population reaches carrying capacity. At this point, there are no surplus individuals to be harvested and yield drops to zero. The maximum sustainable yield is usually higher than the optimum sustainable yield and maximum economic yield

%There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983). This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource. 
%For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis. Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.

%As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing. It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere. Subsidies or external economic considerations such as fuel taxes will also affect the location of an economic Reference Point (e.g. Panayotou 1988).

%The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced. This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g. the tuna fishery for the canning industry.

%The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition. Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points. Even if the actual target F cannot be estimated, in theory, F could be adjusted in increments until the desirable target catch characteristics are achieved.

%In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points. In evaluating investment projects, including resource management, economists discount the future value of any commodity. Discount rates may be in the order of 10\%. In the case of a fishery where the population growth rate does not exceed the discount rate, then a strict application of economic theory would suggest that in the absence of other considerations (such as an economic value placed on recreational use of resources) the whole stock should be harvested now, and the proceeds of their sale invested. Long-lived species with slow growth rates, such as whales, clearly fall into this category. The blatant contradiction between this common economic approach, and the concept of sustainability, constitutes an unresolved paradox (Hilborn and Walters 1992).

>>>>

Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged. The quota constraints were met by a reduction in fishing effort (Fig. \ref{f:effort}). as predicted in general terms by \cite{Ulrich2017}. This led to a reduction in fishing cost and an increase in net revenue over the mid-term because biomasses and yields increased. However, individual quota cobined with the stochastic fisheries catches resulted in lower than intended exploitation rates. The TACs were set so that the total TACs corresponded to the harvest rate that resulted in maximum sustainable yields. The TACs were divided in individual quotas for each fishing vessel. The lower than intended harvest rates were caused by the fact that despite their planning, fishers often exhausted one of the quotas before the other. If discarding is not allowed, then individual fishers must then stop fishing. At the population levels of the two species this resulted in overal cacthes that were lower than the TACs. Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere \cite{Kempf2016, Rindorf2017, Hilborn2015}, but is generall attributed to differences in life-histories among fished species. In this study, the discrepancy between the intended annual quotas at MSY and realized catches is caused only by schochasticity in the system. This stochasticy in catch rates is observed in most fisheries \cite{Sampson1988, VanOostenbrugge2004, Smith1980, Bernasconi2015}. It is caused by the complex migration biology and the effects of environmental factors on fish behaviour. A well functioning quota market would allow vessels to swap excess quota for constraining and thus bring realized catches closer to quotas.

In the case study explored here, setting quotas based on MSY for all species resulted in mixed fisheries that were economically more viable in the long run compared to the unconstrained cased. It should be noted that this particular result may not hold true for mixed fisheries with unavoidable catches or bycatches of very low productive species. Fishing those species at their MSY levels would probably constrain effort to such an extent tat the net revenue will be reduced in the long-term \cite{Prellezo2016a,  Rindorf2017}.

The model aimed to evaluate the biological and economic effects of adopting a major policy change such as the CFP \cite{CFP2013}. Starting from an overfished mixed fishery under open access conditions but with a landing obligation, the response of individual effort allocation and the interannual population dynamics are explicitly modeled, while adopting MSY objectives. The analysis extended earlier approaches that used short-term state-dependent decision-making models (e.g. \cite{Alzorriz2018, Batsleer2016, Poos2010}) by including interannual population dynamics of fish stocks, and a management procedure that mimics the decisions of a management system. Such an MSE framework was thought of as a minimum realistic model \cite{Kell2007, Punt1995}. To keep the outcomes of the model simple, and to show the implications of the fishing effort allocation, the two fish populations modelled here are exactly the same apart from their spatial distribution difference (Fig. \ref{f:distribution}). As a result, if the landing obligation changes selectivity patterns in the short-term as predicted in \cite{Alzorriz2018, Batsleer2016}, the present MSE modeling approach allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling how changes in exploitation rates impact fish populations and yields.

The forecasted effects of fisheries management in mixed fisheries result from spatial heterogeneity in resources and fishers adaptive behaviour. This emphasizes that human behaviour and decision-making are intrinsic parts of conservation and natural resource management and the need for incorporating detailed fishing dynamics in management strategy evaluation. Historically, the focus of management strategy evaluation has been primarily on population dynamics in single and multispecies fisheries \cite{Butterworth2007, Dichmont2006b, Punt1999, Punt2009, Smith1999, Punt2016}. Explicitly incorporating direct effects of human behaviour of those who harvest stocks is equally important \cite{Fulton2011,Milner-Gulland2012, Link2017}, but relatively few studies have attempted this \cite{Andersen2010, Ono2018, Simons2015a}. Our MSE that links a dynamic state variable model for fisher behavior to a multi-species age structured model provides an alternative to those approaches.

The complexity of adding fishers behaviour in a minimum realistic model of mixed fisheries in an MSE framework requires making several simplifying assumptions. The details and possible consequences of these assumptions have been discussed in eg. \cite{Alzorriz2018}, and include full compliance of fishers with the discarding policy, absence of a quota market, monthly choices on fishing, and a lack of interference competition among fishers. Also, the utility function only comprises of decisions based on annual net revenues, ommitting any social dynamics or tradition. These social dynamics and traditions are potentially important \cite{Girardin2017, Schluter2012}, and could be incorporated in future applications. Important assumptions about the dynamics of the population dynamics included negligible natural mortality, constant recruitment, constant growth, and constant migration.

In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annaul quota swaps. We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY \cite{Prellezo2016}. Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries \cite{Alzorriz2018}. The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance. With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Batsleer2013}. In this MSE framework, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations. However, if detection rates are low and the short-term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance.

@@
